---
title: "Microbiome Integrated Reconstruction and Estimation"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![](https://cranlogs.r-pkg.org/badges/mire)](https://cran.r-project.org/package=mire)
<!-- badges: end -->

An integrated framework for microbiome reconstruction from sequencing data. It
leverages tools like Kraken2 for taxonomic classification and combines cell
barcodes, UMIs, and k-mer-based quantification to reconstruct microbial signals.
Designed for both bulk and single-cell sequencing data, the package enables
taxonomic and quantitative profiling of microbial communities.

## Installation
Here, we use `pak` to install package, you can also use `remotes`:

```{r, eval=FALSE}
if (!requireNamespace("pak")) {
    install.packages("pak",
        repos = sprintf(
            "https://r-lib.github.io/p/pak/devel/%s/%s/%s",
            .Platform$pkgType, R.Version()$os, R.Version()$arch
        )
    )
}
```

You can install the development version from
[r-universe](https://wanglabcsu.r-universe.dev/mire) with:

```{r, eval=FALSE}
pak::repo_add("https://wanglabcsu.r-universe.dev")
pak::pak("mire")
```

or from [GitHub](https://github.com/WangLabCSU/mire) with:

```{r, eval=FALSE}
pak::pak("WangLabCSU/mire")
```

You must also install [kraken2](https://github.com/DerrickWood/kraken2/wiki/Manual).

## sessionInfo
```{r}
sessionInfo()
```
