---
title: "Single-cell Analysis of Host-Microbiome Interactions"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

## Installation
Here, we use `pak` to install package, you can also use `remotes`
```{r, eval=FALSE}
if (!requireNamespace("pak")) {
    install.packages("pak",
        repos = sprintf(
            "https://r-lib.github.io/p/pak/devel/%s/%s/%s",
            .Platform$pkgType, R.Version()$os, R.Version()$arch
        )
    )
}
```

`rsami` depends on [polars](https://rpolars.github.io/index.html), you can
install it by:
```{r, eval=FALSE}
pak::repo_add("https://rpolars.r-universe.dev")
pak::pkg_install("polars")
```

You can also install the development version of `polars` from
[Github](https://github.com/pola-rs/r-polars) 

In this way, you must install `rustc` to compile `polars`

You can install `rustc` with `rustup`:
```
curl --proto '=https' --tlsv1.2 -sSf https://sh.rustup.rs | sh
```
Then install the `polars` package:
```{r, eval=FALSE}
pak::pkg_install("pola-rs/r-polars")
```

You can install the development version of `rsahmi` from
[GitHub](https://github.com/Yunuuuu/rsahmi):

```{r, eval=FALSE}
pak::pkg_install("Yunuuuu/rsahmi")
```

You must also install [seqkit](https://bioinf.shenwei.me/seqkit/) and
[kraken2](https://github.com/DerrickWood/kraken2/wiki/Manual). 


## Workflow
```{r}
library(ggplot2)
library(polars)
```
Prepare output directory and input data
```{r, eval=FALSE}
odir <- "rsahmi"
fq1 <-
# For 10x Genomic data, `fq1` only contain barcode and umi, but the official
# didn't give any information for this. In this way, I prefer using `umi-tools` 
# to transform the `umi` into fq2 and then run `rsahmi` with only fq2.
fq2 <- # can be `NULL`
kraken_db <- # specify the kraken database
if (dir.exists(odir)) dir.create(odir, recursive = TRUE)
```

### Taxonomic sequence classifier 
We must run `kraken2` with `--use-names`, and `--report-minimizer-data`,
`classified_out` is not necessary in `rsahmi` package, but should be specified
if you want to use official `SAHMI` script.

```{r, eval=FALSE}
biosys::kraken2(
    fq1 = fq1,
    fq2 = fq2,
    classified_out = "classified.fq",
    biosys::arg("--threads", 10L, format = "%d"),
    biosys::arg("--db", kraken_db),
    "--use-names", "--report-minimizer-data",
    odir = file.path(odir, sample) # you should change sample into your sample name
)$run()
```

### Extract microbiome kraken output
```{r, eval=FALSE}
# you should change sample into your sample name
# 
# this'll extract all taxids specified in `taxon`
# which default to: c("d__Bacteria", "d__Fungi", "d__Viruses")
taxids <- rsahmi::extract_taxids(file.path(odir, sample, "kraken_report.txt"))
# then we extract the kraken2 output for these taxon.
rsahmi::extract_kraken_output(
    file.path(odir, sample, "kraken_output.txt"),
    taxids = taxids,
    dir = file.path(odir, sample)
)
```

### Extract microbiome kraken reads
```{r, eval=FALSE}
# you should change sample into your sample name
rsahmi::extract_kraken_reads(
    kraken_out = file.path(odir, sample, "kraken_microbiome_output.txt"),
    reads = c(fq1, fq2),
    odir = file.path(odir, sample),
    threads = 10L,
    # try to change `seqkit` argument into your seqkit path
    # If `NULL`, the internal will detect it in your `PATH` environment variable
    seqkit = NULL
)
```


**Following workflow will need all samples (multiple), you should run above command for your all samples, then proceed to the subsequent steps.**

### Prepare datasets for signal denoising and taxa counting
```{r, eval=FALSE}
# paths: the output directory of all samples (from above steps)
paths <- fs::dir_ls(odir, type = "directory")
# this is the time-consuming step, so the internal will save the results in `odir`
sahmi_datasets <- lapply(paths, function(dir) {
    rsahmi::prep_dataset(
        fa1 = file.path(dir, "kraken_microbiome_reads.fa"),
        # if you have paired sequence, please also specify `fa2`,
        # !!! Also pay attention to the file name of `fa1` (add suffix `_1`) if
        #     you use paired reads.
        # fa2 = file.path(dir, "kraken_microbiome_reads_2.fa"),
        kraken_report = file.path(dir, "kraken_report.txt"),
        kraken_out = file.path(dir, "kraken_microbiome_output.txt"),
        odir = dir
    )
})
# If you have run above command, you can use following code
# to read the datasets
# sahmi_datasets <- lapply(paths, rsahmi::read_dataset)
```

### Identify contaminants
```{r, eval=FALSE}
truly_microbe <- rsahmi::remove_contaminants(
    file.path(paths, "kraken_report.txt"),
    quantile = 0.99, exclusive = FALSE
)
microbe_for_plot <- attr(truly_microbe, "truly")[
    order(attr(truly_microbe, "pvalue")[attr(truly_microbe, "truly")])
]
microbe_for_plot <- microbe_for_plot[
    !microbe_for_plot %in% attr(truly_microbe, "exclusive")
]

truly_microbe_density <- ggplot(
    truly_microbe$filter(pl$col("taxid")$is_in(microbe_for_plot))$
        to_data_frame(),
    aes(rpmm),
) +
    geom_density(aes(fill = study), alpha = 0.5) +
    scale_x_log10() +
    facet_wrap(facets = vars(taxa), scales = "free") +
    theme(
        strip.clip = "off",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(1, 0),
        legend.justification.inside = c(1, 0)
    )
ggsave("truly_microbe.pdf",
    plot = truly_microbe_density,
    width = 8, height = 7L
)
```

### Sample level signal denoising
```{r, eval=FALSE}
slsd <- rsahmi::slsd(lapply(sahmi_datasets, `[[`, "kreport"))
real_taxids_slsd <- slsd$filter(
    pl$col("r1")$gt(0),
    pl$col("r2")$gt(0),
    pl$col("r3")$gt(0),
    pl$col("p1")$lt(0.05),
    pl$col("p2")$lt(0.05),
    pl$col("p3")$lt(0.05)
)$get_column("taxid")
```

### Filter contaminants
```{r, eval=FALSE}
umi_list <- lapply(sahmi_datasets, function(dataset) {
    # Barcode level signal denoising (barcode k-mer correlation test)
    blsd <- rsahmi::blsd(dataset$kmer)
    real_taxids <- blsd$filter(pl$col("padj")$lt(0.05))$get_column("taxid")
    # only keep taxids pass Sample level signal denoising
    real_taxids <- real_taxids$filter(real_taxids$is_in(real_taxids_slsd))
    # remove contaminants
    real_taxids <- real_taxids$filter(
        real_taxids$is_in(attr(truly_microbe, "truly"))
    )
    # filter UMI data
    dataset$umi$filter(pl$col("taxid")$is_in(real_taxids))
})
```

### Taxa umi counting
```{r, eval=FALSE}
counts <- rsahmi::taxa_counts(umi_list, basename(names(umi_list)))
counts$write_csv(file.path(odir, "taxa_counts.txt"), separator = "\t")
```

## sessionInfo
```{r}
sessionInfo()
```
